---
title: "atac differential"
output: html_document
author: "A. Thomas"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
suppressPackageStartupMessages({
  library(DESeq2)
  library(ggplot2)
  library(pheatmap)
  library(RColorBrewer)
  library(gplots)
  library(ggplot2)
  library(KEGG.db)
  library(GO.db)
  library(goseq)
  library(TxDb.Hsapiens.UCSC.hg19.knownGene)
  library(org.Hs.eg.db)
  library(GenomeInfoDb)
  library(ChIPseeker)
  library(rGREAT)
})
```

#granges from bed3
```{r}
create_granges_frombed3 = function(bed){
  gr1 = read.table(bed,sep="\t",header = F)
  head=c("chr","start","end")
  names(gr1)=head
  gr = GRanges(seqnames=gr1$chr,
               ranges=IRanges(start=gr1$start+1,end=gr1$end),
               strand=rep("*",nrow(gr1)))
  return(gr)
}

```

```{r}
create_grangesnew_frombed3 = function(bed){
  gr1 = bed
  gr = GRanges(seqnames=gr1$seqnames,
               ranges=IRanges(start=gr1$start+1,end=gr1$end),
               strand=rep("*",nrow(gr1)))
  return(gr)
}

```

# barplot
```{r}
barplot <- function(df, showCategory=15){
  #df <- df[with(df, order(ratio, padj, decreasing = c(TRUE, FALSE))),]
  df <- df[with(df, order( padj, decreasing = c(FALSE))),]
  df <- head(df, n=showCategory)
  breaks <- round( c(0, 1/4, 2/4, 3/4, 1) * max(df[['ratio']]) , 2)
  p_plot <- ggplot(df, aes_string(x="term", y="ratio", fill="padj")) + geom_col() +
    scale_y_continuous(expand=c(0, 0), breaks=breaks, limits=c(0, max(df[["ratio"]]+0.05))) +
    scale_x_discrete(name='term') + theme(text=ggplot2::element_text(size=9)) +
    coord_flip() +  theme_bw(base_size=9) + scale_color_gradient(low="blue", high="red") +
    ylab("Ratio of DE genes in Category") +
    xlab("KEGG pathway")
  #scale_fill_continuous(low="#00dbde", high="#FFF94C") + 
  return(p_plot)
}
```

# dotplot
```{r}
dotplot <- function(df, showCategory=15){
  #df <- df[with(df, order(ratio, padj, decreasing = c(TRUE, FALSE))),]
  df <- df[with(df, order( padj, decreasing = c(FALSE))),]
  df <- head(df, n=showCategory)
  d_plot <- ggplot(df, aes_string(x="term", y="ratio", colour="padj",size="numDEInCat")) + 
    geom_point() + scale_color_gradient(low="blue", high="red")+ coord_flip() +
    theme_bw(base_size=9) +
    ylab("Ratio of DE genes in Category") +
    xlab("KEGG pathway")
  #scale_color_gradient(low="#00dbde",   high="#FFF94C") +
  return(d_plot)
}
```

# rGREAT analysis
```{r}
great = function(commonPeaks){
  
seqlevelsStyle(commonPeaks) <- "UCSC"
great_Job <- submitGreatJob(commonPeaks, species = "hg19")
availableCategories(great_Job)
great_ResultTable_motifs = getEnrichmentTables(great_Job, category = "Regulatory Motifs",request_interval=350)
d.motif = as.data.frame(great_ResultTable_motifs$`MSigDB Predicted Promoter Motifs`)
d.motif.filt=d.motif[d.motif$Binom_Raw_PValue<1e-10,]
if(nrow(d.motif.filt)>0){
  data = cbind(d.motif.filt$ID,d.motif.filt$name,d.motif.filt$Binom_Raw_PValue)
  colnames(data)=c("Id","name","Bin pvalue")
  write.table(data,file = paste0("ontology/",outputname,"_MOTIF_Great.tsv"), row.names=F,sep="\t",quote = F)
}
#great_ResultTable_phentype = getEnrichmentTables(great_Job, category = "Phenotype Data and Human Disease")
#great_go =getEnrichmentTables(great_Job, category = "GO")
great_pathway =  getEnrichmentTables(great_Job, category = "Pathway Data")
d=as.data.frame(great_pathway$`PANTHER Pathway`)
d.filt=d[d$Binom_Raw_PValue<0.05,]
if(nrow(d.filt)>0){
  data = cbind(d.filt$ID,d.filt$name,d.filt$Binom_Raw_PValue)
  colnames(data)=c("Id","name","Bin pvalue")
  write.table(data,file = paste0("ontology/",outputname,"_PANTHER_Great.tsv"), row.names=F,sep="\t",quote = F)
}
}
```

# KEGG pathway
```{r}
kegg = function(bed1,bed2,genome,outputname){
  
  create_granges_frombed = function(bed){
    #gr1 = read.table(bed,sep="\t",header = T)
    gr1=bed
    #head=c("chr","start","end")
    #names(gr1)=head
    gr = GRanges(seqnames=gr1$chr,
                 ranges=IRanges(start=gr1$start+1,end=gr1$end),
                 strand=rep("*",nrow(gr1)))
    return(gr)
  }
  
  gr1=create_granges_frombed(bed1)
  gr2=create_granges_frombed(bed2)
  
  allPeaks <- c(gr1,gr2)
  commonPeaks <- reduce(allPeaks)
  Txdb<- TxDb.Hsapiens.UCSC.hg19.knownGene
  hg19_genes<- genes(Txdb)
  promoters= promoters(hg19_genes,upstream = 1000,downstream = 1000)
  allGenes <- unique(unlist(keys(TxDb.Hsapiens.UCSC.hg19.knownGene, "GENEID")))
  length(allGenes)
  # print(paste0("up ",length(gr1)))
  # print(paste0("down ",length(gr2)))
  # print(paste0("combined ",length(commonPeaks)))
  
    xx <- as.list(KEGGPATHID2NAME)
  temp <- cbind(names(xx),unlist(xx))
  addKeggTogoseq <- function(JX,temp){
    for(l in 1:nrow(JX)){
      if(JX[l,1] %in% temp[,1]){
        JX[l,"term"] <- temp[temp[,1] %in% JX[l,1],2]
        JX[l,"ontology"] <- "KEGG"
      }
      
    }
    return(JX)
  }
  
  if(length(gr1)){
      peakAnno.up <- annotatePeak(gr1, tssRegion = c(-1000, 1000), TxDb = TxDb.Hsapiens.UCSC.hg19.knownGene,annoDb = "org.Hs.eg.db")
      output=paste0("annotation/annotate_UP",outputname,".tsv")
      write.table(as.data.frame(peakAnno.up),file = output,sep="\t",quote = F, row.names = F)
      peakAnno_GR.up <- as.GRanges(peakAnno.up)
      genesWithPeakInTSS.up <- unique(peakAnno_GR.up[peakAnno_GR.up$annotation == "Promoter",]$geneId)
      genesWithPeakInTSS1.up = unique(promoters[promoters %over% gr1]$gene_id)
      allGenesForGOseq.up <- as.integer(allGenes %in% genesWithPeakInTSS.up)
      allGenesForGOseq1.up <- as.integer(allGenes %in% genesWithPeakInTSS1.up)
      names(allGenesForGOseq.up) <- allGenes
      names(allGenesForGOseq1.up) <- allGenes
      pwf.up = nullp(allGenesForGOseq.up, "hg19", "knownGene", plot.fit = FALSE)
      pwf1.up = nullp(allGenesForGOseq1.up, "hg19", "knownGene", plot.fit = FALSE)
      go.mf.up<-goseq(pwf.up,genome="hg19","knownGene", test.cats = c("GO:BP","GO:MF"))#, method = "Hypergeometric")#"GO:BP","GO:MF",
      go.mf1.up<-goseq(pwf1.up,genome="hg19","knownGene", test.cats = c("GO:BP","GO:MF"))
      go.mf.up$padj = p.adjust(go.mf.up$over_represented_pvalue, method="BH")
      go.mf.up$ratio <- go.mf.up$numDEInCat/ go.mf.up$numInCat
      go.mf1.up$padj = p.adjust(go.mf1.up$over_represented_pvalue, method="BH")
      go.mf1.up$ratio <- go.mf1.up$numDEInCat / go.mf1.up$numInCat
      
      if(nrow(go.mf.up)>0){
      head(go.mf.up)
      fig=barplot(go.mf.up, showCategory = 15)
      ggsave(plot = fig,filename=paste0("plots/",outputname,"_GO_UP.pdf"))
      }
      if(nrow(go.mf1.up)>0){
      head(go.mf1.up)
      fig=barplot(go.mf1.up, showCategory = 15)
      ggsave(plot = fig,filename=paste0("plots/",outputname,"_hg19prom_GO_UP.pdf"))
      }
      enriched.go.up = go.mf.up$category[p.adjust(go.mf.up$over_represented_pvalue,method="BH")<0.5]
      restemp.mf.up = addKeggTogoseq(go.mf.up,temp)
      restemp.mf1.up = addKeggTogoseq(go.mf1.up,temp)
      go.up = goseq(pwf.up,genome="hg19","knownGene", test.cats = c("KEGG"))#,method = "Hypergeometric")
      go1.up = goseq(pwf1.up,genome="hg19","knownGene", test.cats = c("KEGG"))#,method = "Hypergeometric")
      restemp.up <-addKeggTogoseq(go.up,temp)
      restemp1.up <-addKeggTogoseq(go1.up,temp)
      restemp.up$padj = p.adjust(restemp.up$over_represented_pvalue, method="BH")
      restemp.up$ratio = restemp.up$numDEInCat/restemp.up$numInCat
      
      restemp1.up$padj = p.adjust(restemp1.up$over_represented_pvalue, method="BH")
      restemp1.up$ratio = restemp1.up$numDEInCat/restemp1.up$numInCat
      
      if(nrow(restemp.up)>0){
      head(restemp.up)
      fig=dotplot(restemp.up, showCategory = 15)
      ggsave(plot = fig,filename=paste0("plots/",outputname,"_kEGGPATHWAY_UP.pdf"))
      }
      if(nrow(restemp1.up)>0){
      head(restemp1.up)
      fig=dotplot(restemp1.up, showCategory = 15)
      ggsave(plot = fig,filename=paste0("plots/",outputname,"_kEGGPATHWAY_hg19prom_UP.pdf"))
      }
      print("chipseeker")
      print(head(restemp.up))
      
      print("hg19 annotation")
      print(head(restemp1.up))
      if(nrow(restemp.up)>0)
      {
      write.table(restemp.up,file=paste0("ontology/",outputname,"_UP_Kegg.tsv"), row.names=F,sep="\t",quote = F)
      write.table(restemp1.up,file=paste0("ontology/",outputname,"_UP_hg19Prom_Kegg.tsv"), row.names=F,sep="\t",quote = F)
      }
      
      if(nrow(restemp.mf.up)>0)
      {
      write.table(restemp.mf.up,file=paste0("ontology/",outputname,"_UP_GO_MF.tsv"), row.names=F,sep="\t",quote = F)
      write.table(restemp.mf1.up,file=paste0("ontology/",outputname,"_UP_hg19Prom_GO_MF.tsv"), row.names=F,sep="\t",quote = F)
      }
  }
  
  if(length(gr2)){
    peakAnno.down <- annotatePeak(gr2, tssRegion = c(-1000, 1000), TxDb = TxDb.Hsapiens.UCSC.hg19.knownGene,annoDb = "org.Hs.eg.db")
    output=paste0("annotation/annotate_Down",outputname,".tsv")
    write.table(as.data.frame(peakAnno.down),file = output,sep="\t",quote = F, row.names = F)
    peakAnno_GR.down <- as.GRanges(peakAnno.down)
    genesWithPeakInTSS.down <- unique(peakAnno_GR.down[peakAnno_GR.down$annotation == "Promoter",]$geneId)
    genesWithPeakInTSS1.down = unique(promoters[promoters %over% gr2]$gene_id)
    allGenesForGOseq.down <- as.integer(allGenes %in% genesWithPeakInTSS.down)
    allGenesForGOseq1.down <- as.integer(allGenes %in% genesWithPeakInTSS1.down)
    names(allGenesForGOseq.down) <- allGenes
    names(allGenesForGOseq1.down) <- allGenes
    pwf.down = nullp(allGenesForGOseq.down, "hg19", "knownGene", plot.fit = FALSE)
    pwf1.down = nullp(allGenesForGOseq1.down, "hg19", "knownGene", plot.fit = FALSE)
    go.mf.down<-goseq(pwf.down,genome="hg19","knownGene", test.cats = c("GO:BP","GO:MF"))#, method = "Hypergeometric")#"GO:BP","GO:MF",
    go.mf1.down<-goseq(pwf1.down,genome="hg19","knownGene", test.cats = c("GO:BP","GO:MF"))
    
    go.mf.down$padj = p.adjust(go.mf.down$over_represented_pvalue, method="BH")
    go.mf.down$ratio <- go.mf.down$numDEInCat / go.mf.down$numInCat
    go.mf1.down$padj = p.adjust(go.mf1.down$over_represented_pvalue, method="BH")
    go.mf1.down$ratio <- go.mf1.down$numDEInCat / go.mf1.down$numInCat
    if(nrow(go.mf.down)>0){
    head(go.mf.down)
    fig=barplot(go.mf.down, showCategory = 15)
    ggsave(plot = fig,filename=paste0("plots/",outputname,"_GO_DOWN.pdf"))
    }
    if(nrow(go.mf1.down)>0){
    head(go.mf1.down)
    fig=barplot(go.mf1.down, showCategory = 15)
    ggsave(plot = fig,filename=paste0("plots/",outputname,"_hg19prom_GO_DOWN.pdf"))
    } 
    enriched.go.down = go.mf.down$category[p.adjust(go.mf.down$over_represented_pvalue,method="BH")<0.5]
    restemp.mf.down = addKeggTogoseq(go.mf.down,temp)
    restemp.mf1.down = addKeggTogoseq(go.mf1.down,temp)
    go.down = goseq(pwf.down,genome="hg19","knownGene", test.cats = c("KEGG"))#,method = "Hypergeometric")
    go1.down = goseq(pwf1.down,genome="hg19","knownGene", test.cats = c("KEGG"))#,method = "Hypergeometric")
    
    restemp.down <-addKeggTogoseq(go.down,temp)
    restemp1.down <-addKeggTogoseq(go1.down,temp)
    restemp.down$padj = p.adjust(restemp.down$over_represented_pvalue, method="BH")
    restemp.down$ratio = restemp.down$numDEInCat/restemp.down$numInCat
    
    restemp1.down$padj = p.adjust(restemp1.down$over_represented_pvalue, method="BH")
    restemp1.down$ratio = restemp1.down$numDEInCat/restemp1.down$numInCat
    if(nrow(restemp.down)>0){
    head(restemp.down)
    fig=dotplot(restemp.down, showCategory = 15)
    ggsave(plot = fig,filename=paste0("plots/",outputname,"_kEGGPATHWAY_DOWN.pdf"))
    }
    if(nrow(restemp1.down)>0){
    head(restemp1.down)
    fig=dotplot(restemp1.down, showCategory = 15)
    ggsave(plot = fig,filename=paste0("plots/",outputname,"_kEGGPATHWAY_hg19prom_DOWN.pdf"))
    }
    print("chipseeker")
    print(head(restemp.down))
    print("hg19 annotation")
    print(head(restemp1.down))
    print(nrow(restemp.down))
    if(nrow(restemp.down)>0)
    {
    write.table(restemp.down,file=paste0("ontology/",outputname,"_Down_Kegg.tsv"), row.names=F,sep="\t",quote = F)
    write.table(restemp1.down,file=paste0("ontology/",outputname,"_Down_hg19Prom_Kegg.tsv"), row.names=F,sep="\t",quote = F)
    }
    
    if(nrow(restemp.mf.down)>0)
    {
    write.table(restemp.mf.down,file=paste0("ontology/",outputname,"_Down_GO_MF.tsv"), row.names=F,sep="\t",quote = F)
    write.table(restemp.mf1.down,file=paste0("ontology/",outputname,"_Down_hg19Prom_GO_MF.tsv"), row.names=F,sep="\t",quote = F)
    }
  }
  
  #Call Great Function
  great(commonPeaks)
}

```

#Write outputfiles
```{r}

createoutput = function(res, rld1, outputname){
  
  rlogcount <- assay(rld1)
  sampleDists <- as.matrix(dist(t(rlogcount)))
  showcols <- brewer.pal(8, "Set1")[1:length(unique(colData(rld1)$cellline))]
  a=data.frame(name=colData(rld1)$cellline)
  hmcol <- colorRampPalette(brewer.pal(9, "GnBu"))(100)
  pdf(file = paste0("plots/MDS_PCA_Heatmap/",outputname,"_atac_sampledist.pdf"))
  heatmap.2(as.matrix(sampleDists), key=F, trace="none",
            col=hmcol,
                ColSideColors=showcols[factor(a$name)], 
                RowSideColors=showcols[factor(a$name)],
            margin=c(10, 10))
  dev.off()
  test_cellline <- res
  
  #differential regions
  up_fdr <- test_cellline[!is.na(test_cellline$padj) & test_cellline$padj <= 0.05  & test_cellline$log2FoldChange > 0]
  down_fdr <- test_cellline[test_cellline$padj <= 0.05 & !is.na(test_cellline$padj) & test_cellline$log2FoldChange < 0]
  
  up_fdr.log2 <- test_cellline[test_cellline$padj < 0.05 & !is.na(test_cellline$padj) & test_cellline$log2FoldChange >= 1]
  down_fdr.log2 <- test_cellline[test_cellline$padj < 0.05 & !is.na(test_cellline$padj) & test_cellline$log2FoldChange <= -1] 
  up_fdr_0.1 <- test_cellline[test_cellline$padj <= 0.1 & !is.na(test_cellline$padj) & test_cellline$log2FoldChange > 0]
  down_fdr_0.1 <- test_cellline[test_cellline$padj <= 0.1 & !is.na(test_cellline$padj) & test_cellline$log2FoldChange < 0]
  
  up_fdr.log2_0.1 <- test_cellline[test_cellline$padj <= 0.1 & !is.na(test_cellline$padj) & test_cellline$log2FoldChange >= 1]
  down_fdr.log2_0.1 <- test_cellline[test_cellline$padj <= 0.1 & !is.na(test_cellline$padj) & test_cellline$log2FoldChange <= -1] 
  cat("#####\n No. of Differential regions identified:\nfdr_up\n")
  print(length(up_fdr))
  print("fdr_down")
  print(length(down_fdr))
  
  #function to create data frame for output generation
  create_output =function(df,flag){
    if(flag==1) df=df[order(df$padj)]
    if(flag==2) df=df[order(df$pvalue)]
    gr1 <- data.frame(chr=seqnames(df),
                      start=start(df)-1,
                      end=end(df),
                      log2Foldchange=df$log2FoldChange,
                      pvalue=df$pvalue,
                      padj=df$padj)  
    
  }

  if(length(up_fdr)){
    up.regions=create_output(up_fdr,1)
    write.table(up.regions, file=paste0("fdr0.05/",outputname,"_DESEQ2_up_fdr0.05.bed"), quote=F, sep="\t", row.names=F, col.names=T)
  }
  if(length(down_fdr))
  {
    down.regions=create_output(down_fdr,1)
    write.table(down.regions,file=paste0("fdr0.05/",outputname,"_DESEQ2_down_fdr0.05.bed"), quote=F, sep="\t", row.names=F, col.names=T)
  }
  if(length(up_fdr.log2)){
    up.regions=create_output(up_fdr.log2,1)
    write.table(up.regions, file=paste0("fdr0.05/fold2/",outputname,"_DESEQ2.fold2_up_fdr0.05.bed"), quote=F, sep="\t", row.names=F, col.names=T)
  }
  if(length(down_fdr.log2))
  {
    down.regions=create_output(down_fdr.log2,1)
    write.table(down.regions,file=paste0("fdr0.05/fold2/",outputname,"_DESEQ2.fold2_down_fdr0.05.bed"), quote=F, sep="\t", row.names=F, col.names=T)
  }
  
    if(length(up_fdr_0.1)){
    up.regions=create_output(up_fdr_0.1,1)
    write.table(up.regions, file=paste0("fdr0.1/",outputname,"_DESEQ2_up_fdr0.1.bed"), quote=F, sep="\t", row.names=F, col.names=T)
    }
  if(length(down_fdr_0.1))
  {
    down.regions=create_output(down_fdr_0.1,1)
    write.table(down.regions,file=paste0("fdr0.1/",outputname,"_DESEQ2_down_fdr0.1.bed"), quote=F, sep="\t", row.names=F, col.names=T)
  }
  if(length(up_fdr.log2_0.1)){
    up.regions=create_output(up_fdr.log2_0.1,1)
    write.table(up.regions, file=paste0("fdr0.1/fold2/",outputname,"_DESEQ2.fold2_up_fdr0.1.bed"), quote=F, sep="\t", row.names=F, col.names=T)
  }
  if(length(down_fdr.log2_0.1))
  {
    down.regions=create_output(down_fdr.log2_0.1,1)
    write.table(down.regions,file=paste0("fdr0.1/fold2/",outputname,"_DESEQ2.fold2_down_fdr0.1.bed"), quote=F, sep="\t", row.names=F, col.names=T)
  }
  
  #Ontology & pathway analysis
  if(length(up_fdr)+length(down_fdr) > 10){
    kegg(create_output(up_fdr,1),create_output(down_fdr,1),"hg19",outputname)
  }
  
  up <- test_cellline[test_cellline$log2FoldChange > 0 & !is.na(test_cellline$pvalue)]
  top_up = create_output(up,2)
  down = test_cellline[test_cellline$log2FoldChange < 0 & !is.na(test_cellline$pvalue)]
  top_down = create_output(down,2)
  
  write.table(top_up, file=paste0("otheroutput/",outputname,"_DESEQ2_up_sortpval.bed"), quote=F, sep="\t", row.names=F, col.names=T)
  write.table(top_up, file=paste0("otheroutput/",outputname,"_DESEQ2_up_sortpval.bed"), quote=F, sep="\t", row.names=F, col.names=F)
  write.table(head(top_up,n=100L), file=paste0("otheroutput/",outputname,"_DESEQ2_up_sortpval_top100.bed"), 
              quote=F, sep="\t", row.names=F, col.names=F)
  
  write.table(top_down, file=paste0("otheroutput/",outputname,"_DESEQ2_down_sortpval.bed"), quote=F, sep="\t", row.names=F, col.names=T)
  write.table(top_down, file=paste0("otheroutput/",outputname,"_DESEQ2_down_sortpval.bed"), quote=F, sep="\t", row.names=F, col.names=F)
  write.table(head(top_down,n=100L), file=paste0("otheroutput/",outputname,"_DESEQ2_down_sortpval_top100.bed"),
              quote=F, sep="\t", row.names=F, col.names=F)

  data <- plotPCA(rld1, intgroup="cellline", returnData=TRUE)
  percentVar <- round(100 * attr(data, "percentVar"))
  #pdf(file =  paste0("plots/",output.name,"_RNA_PCA.pdf"))
  ggplot(data, aes(PC1, PC2,label=rownames(colData(rld1))))+ theme_bw() +
    theme(axis.line = element_line(colour = "black"),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          #panel.border = element_blank(),
          panel.background = element_blank()) +
   geom_text(col=showcols[factor(a$name)],size=2.5,check_overlap = FALSE)+
    xlab(paste0("PC1: ",percentVar[1],"% variance")) +
    ylab(paste0("PC2: ",percentVar[2],"% variance"))
  ggsave(filename =paste0("plots/MDS_PCA_Heatmap/",outputname,"_atac_PCA.pdf") )
}

```

# create directories
```{r}
mainDir=getwd()
subDir="plots"
dir.create(file.path(mainDir,subDir), showWarnings = FALSE)
subDir="plots/MDS_PCA_Heatmap"
dir.create(file.path(mainDir,subDir), showWarnings = FALSE)
subDir="fdr0.05"
dir.create(file.path(mainDir,subDir), showWarnings = FALSE)
subDir="fdr0.05/fold2"
dir.create(file.path(mainDir,subDir), showWarnings = FALSE)
subDir="fdr0.1"
dir.create(file.path(mainDir,subDir), showWarnings = FALSE)
subDir="fdr0.1/fold2"
dir.create(file.path(mainDir,subDir), showWarnings = FALSE)
dir.create(file.path(mainDir,subDir), showWarnings = FALSE)
subDir="otheroutput"
dir.create(file.path(mainDir,subDir), showWarnings = FALSE)
subDir="countdata"
dir.create(file.path(mainDir,subDir), showWarnings = FALSE)
subDir="annotation"
dir.create(file.path(mainDir,subDir), showWarnings = FALSE)
subDir="ontology"
dir.create(file.path(mainDir,subDir), showWarnings = FALSE)
```

# Read counts table
```{r}
path="data/"
df = read.table(paste0(path,"atac_rawcounts.tsv"),sep="\t",header=T,row.names = 1)
head(df[,1:5])
dim(df)
```


# Read metainfo
```{r}
design.matrix = read.table(paste0(path,"atac_design_matrix.tsv"),sep="\t",header = T,row.names = 1)
design.matrix = as.data.frame(design.matrix[match(colnames(df),rownames(design.matrix)),])
design.matrix
```

#write design matrix
```{r}
write.table(df,file = "atac_rawcounts.tsv",sep = "\t",quote = F, col.names = NA)
write.table(design.matrix,file = "atac_design_matrix.tsv",sep = "\t", quote = F, col.names = NA)
```

# Granges for all regions
```{r}
bedfile="atac_peaks.merged.bed"
totalwindows=create_granges_frombed3(paste0(path,bedfile))
length(totalwindows)
df_new <- df1[ rowSums(df1) > 100,  ]
paste0(nrow(df_new),"*",ncol(df_new))
totalwindows_new = create_grangesnew_frombed3(data.frame(totalwindows)[as.logical(rowSums(df1) >100),])
nrow(df_new)
length(totalwindows_new)
```


# DESEQ2 call parent vs non parent
```{r}
dds  = DESeqDataSetFromMatrix(countData= df_new, colData= design.matrix, ~ culture + dissociation_treatment + celllinetargetorgan , rowRanges=totalwindows_new)
dds =  DESeq(dds)#, minReplicatesForReplace=Inf)
# dds <- estimateSizeFactors(dds)
# dds <- estimateDispersions(dds)
# dds <- nbinomWaldTest(dds, maxit=500)
rld <-varianceStabilizingTransformation(dds,blind = TRUE)
ret <- counts(dds, normalized=T) 
write.table(ret,file = "countdata/atac_countdata_DESEQ2_norm.tsv",sep="\t",row.names = TRUE,quote = FALSE,col.names = NA)
write.table(assay(rld),file = "countdata/atac_countdata_RLD_norm.tsv",sep="\t",row.names = TRUE,quote = FALSE,col.names = NA)
```

#save image
```{r}
resultsNames(dds)
#save.image(file = "atacdifferential_uniquedissociation.RData")
```

# 4.1 Brx68brai vs Brx68H
```{r}
outputname = "Comparison4.1_Brx68brain-vs-68parents"
res = results(dds, alpha=0.05, contrast = c("celllinetargetorgan","Brx68brai","Brx68H"),format = "GRanges")
keep1 = rowSums(counts(dds[,dds$celllinetargetorgan %in% c("Brx68brai")]) >= 10) >= 1
keep2 = rowSums(counts(dds[,dds$celllinetargetorgan %in% c("Brx68H")]) >= 10) >= 2
res.sub = res[keep1 & keep2,]
summary(res.sub)
rld.sub = rld
print(outputname)
colData(rld.sub)
createoutput(res.sub,rld.sub, outputname)
```

# 4.5 Brx68lung vs Brx68H
```{r}
outputname = "Comparison4.5_Brx68lung-vs-68parents"
res = results(dds, alpha=0.05, contrast = c("celllinetargetorgan","Brx68lung","Brx68H"),format = "GRanges")
keep1 = rowSums(counts(dds[,dds$celllinetargetorgan %in% c("Brx68lung")]) >= 10) >= 1
keep2 = rowSums(counts(dds[,dds$celllinetargetorgan %in% c("Brx68H")]) >= 10) >= 2
res.sub = res[keep1 & keep2,]
summary(res.sub)
rld.sub = rld
print(outputname)
colData(rld.sub)
createoutput(res.sub,rld.sub, outputname)
```

# 1.1 Brx07lung vs Brx07H
```{r}
outputname = "Comparison1.1_Brx07lung-vs-07parents"
res = results(dds, alpha=0.05, contrast = c("celllinetargetorgan","Brx07lung","Brx07H"),format = "GRanges")
keep1 = rowSums(counts(dds[,dds$celllinetargetorgan %in% c("Brx07lung")]) >= 10) >= 2
keep2 = rowSums(counts(dds[,dds$celllinetargetorgan %in% c("Brx07H")]) >= 10) >= 2
res.sub = res[keep1 & keep2,]
summary(res.sub)
rld.sub = rld
print(outputname)
colData(rld.sub)
createoutput(res.sub,rld.sub, outputname)
```

# 1.2 Brx07ovar vs Brx07H
```{r}
outputname = "Comparison1.2_Brx07ovar-vs-07parents"
res = results(dds, alpha=0.05, contrast = c("celllinetargetorgan","Brx07ovar","Brx07H"),format = "GRanges")
keep1 = rowSums(counts(dds[,dds$celllinetargetorgan %in% c("Brx07ovar")]) >= 10) >= 2
keep2 = rowSums(counts(dds[,dds$celllinetargetorgan %in% c("Brx07H")]) >= 10) >= 2
res.sub = res[keep1 & keep2,]
summary(res.sub)
rld.sub = rld
print(outputname)
colData(rld.sub)
createoutput(res.sub,rld.sub, outputname)
```

# 2.1 Brx42brai vs Brx07H
```{r}
outputname = "Comparison2.1_Brx42brai-vs-42parents"
res = results(dds, alpha=0.05, contrast = c("celllinetargetorgan","Brx42brai","Brx42H"),format = "GRanges")
keep1 = rowSums(counts(dds[,dds$celllinetargetorgan %in% c("Brx42brai")]) >= 10) >= 2
keep2 = rowSums(counts(dds[,dds$celllinetargetorgan %in% c("Brx42H")]) >= 10) >= 2
res.sub = res[keep1 & keep2,]
summary(res.sub)
rld.sub = rld
print(outputname)
colData(rld.sub)
createoutput(res.sub,rld.sub, outputname)
```


# 3.1 Brx50brai vs Brx07H
```{r}
outputname = "Comparison3.1_Brx50brai-vs-50parents"
res = results(dds, alpha=0.05, contrast = c("celllinetargetorgan","Brx50brai","Brx50H"),format = "GRanges")
keep1 = rowSums(counts(dds[,dds$celllinetargetorgan %in% c("Brx50brai")]) >= 10) >= 2
keep2 = rowSums(counts(dds[,dds$celllinetargetorgan %in% c("Brx50H")]) >= 10) >= 2
res.sub = res[keep1 & keep2,]
summary(res.sub)
rld.sub = rld
print(outputname)
colData(rld.sub)
createoutput(res.sub,rld.sub, outputname)
```

# 3.2 Brx50bone vs Brx07H
```{r}
outputname = "Comparison3.2_Brx50bone-vs-42parents"
res = results(dds, alpha=0.05, contrast = c("celllinetargetorgan","Brx50bone","Brx50H"),format = "GRanges")
keep1 = rowSums(counts(dds[,dds$celllinetargetorgan %in% c("Brx50bone")]) >= 10) >= 2
keep2 = rowSums(counts(dds[,dds$celllinetargetorgan %in% c("Brx50H")]) >= 10) >= 2
res.sub = res[keep1 & keep2,]
summary(res.sub)
rld.sub = rld
print(outputname)
colData(rld.sub)
createoutput(res.sub,rld.sub, outputname)
```


# ComparisonA.mets-vs-parents weights
```{r}
wt=data.frame(a=resultsNames(dds),b=c(0,0,0,0,0,1/9,1/9,1/9,1/9,-1/4,1/9,1/9,-1/4,1/9,1/9,-1/4,1/9))
wt
sum(wt$b)
```


# mets-vs-parents
```{r}
outputname = "ComparisonA_mets-vs-parent"
res = results(dds, alpha=0.05, contrast = c(0,0,0,0,0,1/9,1/9,1/9,1/9,-1/4,1/9,1/9,-1/4,1/9,1/9,-1/4,1/9),format = "GRanges")
keep1 = rowSums(counts(dds[,dds$celllinetargetorgan %in% c("Brx50H","Brx68H","Brx07H","Brx42H")]) >= 10) >= 4
keep2 = rowSums(counts(dds[,dds$celllinetargetorgan %in% c("Brx07lung","Brx07ovar","Brx07lynd", "Brx07kidn","Brx42brai","Brx42ovar","Brx50brai","Brx50bone", "Brx68lung","Brx68bone","Brx68brai")]) >= 10) >= 10
res.sub = res[keep1 & keep2,]
 summary(res.sub)
rld.sub = rld
print(outputname)
createoutput(res.sub,rld.sub, outputname)
#ret <- counts(dds, normalized=T) 
# write.table(ret, paste0("countdata/",outputname,"countdata_norm.txt"),sep="\t",row.names = TRUE,quote = FALSE,col.names = NA)
# createoutput(res.sub, outputname)
# plotfigures(rld.sub)
# kegg(res.sub, outputname,"hg19")
# plotheatmap(res.sub,rld1,20,outputname,"parenttype")
```

# ComparisonB_Brx07allmets-vs-07parents
```{r}
wt=data.frame(a=resultsNames(dds),b=c(0,0,0,0,0,1/3,1/3,1/3,0,0,0,0,0,0,0,0,0))
wt
sum(wt$b)
```

# ComparisonB_Brx07allmets-vs-07parents
```{r}
outputname = "ComparisonB_Brx07allmets-vs-07parents"
res = results(dds, alpha=0.05, contrast = c(0,0,0,0,0,1/3,1/3,1/3,0,0,0,0,0,0,0,0,0),format = "GRanges")
keep1 = rowSums(counts(dds[,dds$celllinetargetorgan %in% c("Brx07lynd","Brx07lung","Brx07ovar")]) >= 10) >= 2
keep2 = rowSums(counts(dds[,dds$celllinetargetorgan %in% c("Brx07H")]) >= 10) >= 2
res.sub = res[keep1 & keep2,]
summary(res.sub)
rld.sub = rld
print(outputname)
colData(rld.sub)
createoutput(res.sub,rld.sub, outputname)
```



# ComparisonC_Brx42allmets-vs-42parents
```{r}
wt=data.frame(a=resultsNames(dds),b=c(0,0,0,0,0,0,0,0,1,-1,0,0,0,0,0,0,0))
wt
sum(wt$b)
```

```{r}
outputname = "ComparisonC_Brx42allmets-vs-42parents"
res = results(dds, alpha=0.05, contrast = c(0,0,0,0,0,0,0,0,1,-1,0,0,0,0,0,0,0),format = "GRanges")
keep1 = rowSums(counts(dds[,dds$celllinetargetorgan %in% c("Brx42ovar","Brx42brai")]) >= 10) >= 2
keep2 = rowSums(counts(dds[,dds$celllinetargetorgan %in% c("Brx42H")]) >= 10) >= 2
res.sub = res[keep1 & keep2,]
summary(res.sub)
rld.sub = rld
print(outputname)
createoutput(res.sub,rld.sub, outputname)
```

# ComparisonD_Brx50allmets-vs-50parents
```{r}
wt=data.frame(a=resultsNames(dds),b=c(0,0,0,0,0,0,0,0,0,0,0.5,0.5,-1,0,0,0,0))
wt
sum(wt$b)
```

```{r}
outputname = "ComparisonD_Brx50allmets-vs-50parents"
res = results(dds, alpha=0.05, contrast =c(0,0,0,0,0,0,0,0,0,0,0.5,0.5,-1,0,0,0,0),format = "GRanges")
keep1 = rowSums(counts(dds[,dds$celllinetargetorgan %in% c("Brx50brai","Brx50bone")]) >= 10) >= 2
keep2 = rowSums(counts(dds[,dds$celllinetargetorgan %in% c("Brx50H")]) >= 10) >= 2
res.sub = res[keep1 & keep2,]
summary(res.sub)
rld.sub = rld
print(outputname)
createoutput(res.sub,rld.sub, outputname)
```



# ComparisonE_Brx68allmets-vs-68parents
```{r}
wt=data.frame(a=resultsNames(dds),b=c(0,0,0,0,0,0,0,0,0,0,0,0,0,1/3,1/3,-1,1/3))
wt
sum(wt$b)
```

```{r}
outputname = "ComparisonE_Brx68allmets-vs-68parents"
res = results(dds, alpha=0.05, contrast = c(0,0,0,0,0,0,0,0,0,0,0,0,0,1/3,1/3,-1,1/3),format = "GRanges")
keep1 = rowSums(counts(dds[,dds$celllinetargetorgan %in% c("Brx68bone","Brx68brai","Brx68lung")]) >= 10) >= 2
keep2 = rowSums(counts(dds[,dds$celllinetargetorgan %in% c("Brx68H")]) >= 10) >= 2
res.sub = res[keep1 & keep2,]
summary(res.sub)
rld.sub = rld
print(outputname)
colData(rld.sub)
createoutput(res.sub,rld.sub, outputname)
```

```{r}
resultsNames(dds)
```

#brain vs other mets weights
```{r}
wt=data.frame(a=resultsNames(dds),b=c(0,0,0,0,0,-1/6,-1/6,-1/6,1/3,0,-1/6,1/3,0,-1/6,1/3,0,-1/6))
wt
sum(wt$b)
```

```{r}
outputname = "ComparisonI_brain-vs-othermets"
res = results(dds, alpha=0.05, contrast = c(0,0,0,0,0,-1/6,-1/6,-1/6,1/3,0,-1/6,1/3,0,-1/6,1/3,0,-1/6),format = "GRanges")
keep1 = rowSums(counts(dds[,dds$celllinetargetorgan %in% c("Brx42brai","Brx50brai","Brx68brai")]) >= 10) >= 2
keep2 = rowSums(counts(dds[,dds$celllinetargetorgan %in% c("Brx07lung","Brx07ovar","Brx07kidn","Brx42ovar","Brx50bone", "Brx68lung","Brx68bone")]) >= 10) >= 2
res.sub = res[keep1 & keep2,]
summary(res.sub)
rld.sub = rld
print(outputname)
colData(rld.sub)
createoutput(res.sub,rld.sub, outputname)
```

#lung vs other mets weights
```{r}
wt=data.frame(a=resultsNames(dds),b=c(0,0,0,0,0,-1/7,1/2,-1/7,-1/7,0,-1/7,-1/7,0,-1/7,-1/7,0,1/2))
wt
sum(wt$b)
```

```{r}
outputname = "ComparisonII_lung-vs-othermets"
res = results(dds, alpha=0.05, contrast = c(0,0,0,0,0,-1/7,1/2,-1/7,-1/7,0,-1/7,-1/7,0,-1/7,-1/7,0,1/2),format = "GRanges")
keep1 = rowSums(counts(dds[,dds$celllinetargetorgan %in% c("Brx07lung","Brx68lung")]) >= 10) >= 2
keep2 = rowSums(counts(dds[,dds$celllinetargetorgan %in% c("Brx42brai","Brx50brai","Brx68brai","Brx07ovar","Brx07kidn","Brx42ovar","Brx50bone", "Brx68lung","Brx68bone")]) >= 10) >= 10
res.sub = res[keep1 & keep2,]
summary(res.sub)
rld.sub = rld
print(outputname)
colData(rld.sub)
createoutput(res.sub,rld.sub, outputname)
```

#bone vs other mets weights
```{r}
wt=data.frame(a=resultsNames(dds),b=c(0,0,0,0,0,-1/7,-1/7,-1/7,-1/7,0,1/2,-1/7,0,1/2,-1/7,0,-1/7))
wt
sum(wt$b)
```


```{r}
outputname = "ComparisonIII_bone-vs-othermets"
res = results(dds, alpha=0.05, contrast = c(0,0,0,0,0,-1/7,-1/7,-1/7,-1/7,0,1/2,-1/7,0,1/2,-1/7,0,-1/7),format = "GRanges")
keep1 = rowSums(counts(dds[,dds$celllinetargetorgan %in% c("Brx50bone","Brx68bone")]) >= 10) >= 2
keep2 = rowSums(counts(dds[,dds$celllinetargetorgan %in% c("Brx07lung","Brx68lung","Brx42brai","Brx50brai","Brx68brai","Brx07ovar","Brx07kidn","Brx42ovar","Brx68lung")]) >= 10) >= 10
res.sub = res[keep1 & keep2,]
summary(res.sub)
rld.sub = rld
print(outputname)
colData(rld.sub)
createoutput(res.sub,rld.sub, outputname)
```

#ovary vs other mets weights
```{r}
wt=data.frame(a=resultsNames(dds),b=c(0,0,0,0,0,-1/8,-1/8,1,-1/8,0,-1/8,-1/8,0,-1/8,-1/8,0,-1/8))
wt
sum(wt$b)
```

```{r}
outputname = "ComparisonIV_ovary-vs-othermets"
res = results(dds, alpha=0.05, contrast = c(0,0,0,0,0,-1/8,-1/8,1,-1/8,0,-1/8,-1/8,0,-1/8,-1/8,0,-1/8),format = "GRanges")
keep1 = rowSums(counts(dds[,dds$celllinetargetorgan %in% c("Brx07ovar","Brx42ovar")]) >= 10) >= 2
keep2 = rowSums(counts(dds[,dds$celllinetargetorgan %in% c("Brx07lung","Brx68lung","Brx42brai","Brx50brai","Brx68brai","Brx07kidn","Brx68lung","Brx50bone","Brx68bone")]) >= 10) >= 10
res.sub = res[keep1 & keep2,]
summary(res.sub)
rld.sub = rld
print(outputname)
colData(rld.sub)
createoutput(res.sub,rld.sub, outputname)
```

```{r}
resultsNames(dds)
```


# 4.2 Brx68bone vs Brx68H
```{r}
outputname = "Comparison4.2_Brx68bone-vs-68parents"
res = results(dds, alpha=0.05, contrast = c("celllinetargetorgan","Brx68bone","Brx68H"),format = "GRanges")
keep1 = rowSums(counts(dds[,dds$celllinetargetorgan %in% c("Brx68bone")]) >= 10) >= 2
keep2 = rowSums(counts(dds[,dds$celllinetargetorgan %in% c("Brx68H")]) >= 10) >= 2
res.sub = res[keep1 & keep2,]
summary(res.sub)
rld.sub = rld
print(outputname)
colData(rld.sub)
createoutput(res.sub,rld.sub, outputname)
```

```{r}
resultsNames(dds)
```

#4.3 Brx68bonemets vs Brx68othermets
```{r}
wt=data.frame(a=resultsNames(dds),b=c(0,0,0,0,0,0,0,0,0,0,0,0,0,1,-1/2,0,-1/2))
wt
sum(wt$b)
```



# 4.3 Brx68bonemets vs Brx68othermets
```{r}
outputname = "Comparison4.3_Brx68bonemets-vs-68othermets"
res = results(dds, alpha=0.05, contrast = c(0,0,0,0,0,0,0,0,0,0,0,0,0,1,-1/2,0,-1/2),format = "GRanges")
#keep1 = rowSums(counts(dds[,dds$celllinetargetorgan %in% c("Brx68bone")]) >= 10) >= 2
#keep2 = rowSums(counts(dds[,dds$celllinetargetorgan %in% c("Brx68brai,Brx68lung")]) >= 10) >= 1
res.sub = res#[keep1 & keep2,]
summary(res.sub)
rld.sub = rld
print(outputname)
colData(rld.sub)
createoutput(res.sub,rld.sub, outputname)
```


# 4.4 Brx68bonemetsplusparents vs Brx68othermets
```{r}
wt=data.frame(a=resultsNames(dds),b=c(0,0,0,0,0,0,0,0,0,0,0,0,0,1/2,-1/2,1/2,-1/2))
wt
sum(wt$b)
```

# 4.4 Brx68bonemetsplusparents vs Brx68othermets
```{r}
outputname = "Comparison4.4_Brx68bonemetsplusparents-vs-68othermets"
res = results(dds, alpha=0.05, contrast = c(0,0,0,0,0,0,0,0,0,0,0,0,0,1/2,-1/2,1/2,-1/2),format = "GRanges")
#keep1 = rowSums(counts(dds[,dds$celllinetargetorgan %in% c("Brx68bone,Brx68H")]) >= 10) >= 2
#keep2 = rowSums(counts(dds[,dds$celllinetargetorgan %in% c("Brx68brai,Brx68lung")]) >= 10) >= 1
res.sub = res#[keep1 & keep2,]
summary(res.sub)
rld.sub = rld
print(outputname)
colData(rld.sub)
createoutput(res.sub,rld.sub, outputname)
```


# ComparisonF_ 4250brainmetsplusparents-vs-0768mets
```{r}
wt=data.frame(a=resultsNames(dds),b=c(0,0,0,0,0,-1/6,-1/6,-1/6,1/4,1/4,0,1/4,1/4,-1/6,0,-1/6,-1/6))
wt
sum(wt$b)
```

# F.4250brainmetsplusparents-vs-0768mets
```{r}
outputname = "ComparisonF_4250brainmetsplusparents-vs-0768mets"
res = results(dds, alpha=0.05, contrast = c(0,0,0,0,0,-1/6,-1/6,-1/6,1/4,1/4,0,1/4,1/4,-1/6,0,-1/6,-1/6),format = "GRanges")
#keep1 = rowSums(counts(dds[,dds$celllinetargetorgan %in% c("Brx68bone")]) >= 10) >= 2
#keep2 = rowSums(counts(dds[,dds$celllinetargetorgan %in% c("Brx68brai,Brx68lung")]) >= 10) >= 1
res.sub = res#[keep1 & keep2,]
summary(res.sub)
rld.sub = rld
print(outputname)
colData(rld.sub)
createoutput(res.sub,rld.sub, outputname)
```